Pushdown Control-Flow Analysis via Refunctionalization (Functional Pearl)

### 2018-03-15 paper.tex
writing

In fact, as later we will eliminate the stack and transform it to continuations
that are represented by higher-order functions, it does not matter what kind of 
AAM we start from. 
Because transforming to higher-order functions forces us to sequentialize 
the order of abstract evaluation.
We choose to start from AAM with unbounded-stack just because it has a 
straightforward correspondence of pushdown semantics to refunctionalized AAM.

@@

fix point cache

@@

\subsection{Higher-Order Controls?}

TODO: analyze programs with call/cc, shift, reset.

@@

\section{Perspective}

\subsection{Defunctionalization}

specify higher-order definitional interpreters using first-order means (cite)

\subsection{Refunctionalization}

%\section{Detour: Shiver's \textit{k}-CFA?}

@@

\textbf{Pushdown Control-flow Analysis}
In recent years, there are significant efforts \cite{vardoulakis2010cfa2, earl2012introspective, 
gilray2016pushdown, johnson2015abstracting} to achieve precise call/return 
match based on small-step AAM which is what we described in section \ref{aam}.

CFA2 is the first solution that solves the return-flows problem\cite{vardoulakis2010cfa2}.
But CFA2 has several drawbacks: it works on only continuation-passing style programs, 
and not supports polyvariant analysis, in addition with an exponential time complexity.
TODO: summarize PDCFA, AAC

\citeauthor{gilray2016pushdown} proposed a polyvariant address allocator for continuations
under small-step abstracting abstract machines that is both simple to implement and
computational inexpensive. 
Based on the AAM we presented in Section \ref{aam}, the only change in code is
not only keeping track of the entry-point expression of callee, but also holding the target environment
when allocating a continuation address.
The target environment is the binding environment extended from closure's environment
(namely @val new_env = env_c + (v -> baddr)@). All the other part of the code remain the same.
The following code shows the implementation:

@@

TODO: explain the reason why it works.

Notably, this change only causes constant-factor increase in time complexity of 
the analysis if the store is widened.

TOOD: why only needs the target env. correspondence?

\textbf{Abstracting Definitional Interpreters.}
On the other side of small-step AAM, with the insight from Reynold's seminal paper 
\emph{Definitional interpreters for higher-order programming languages} \cite{reynolds1972definitional}
that the defined language will inherit properties from the defining language, 
\citeauthor{darais2017abstracting} construct abstracting 
definitional interpreters that automatically inherit the pushdown control-flow 
property from its defining language.

\citeauthor{darais2017abstracting}'s abstracting definitional interpreters is written
in monadic style.

composable.
TODO: connection to this paper.
TODO: although they establish pushdown control-flow analysis in different ways, 
is there a correspondence between different methods?

>>>>

In fact, as later we will eliminate the stack and transform it to continuations
that are represented by higher-order functions, it does not matter what kind of 
AAM we start from. 
Because transforming to higher-order functions forces us to sequentialize 
the order of abstract evaluation.
We choose to start from AAM with unbounded-stack just because it has a 
straightforward correspondence of pushdown semantics to the refunctionalized AAM.

@@

We will apply the fixed-point caching algorithm as described in
\citeauthor{darais2017abstracting}'s abstract definitional interpreter.

@@

@@

%\section{Perspective}
%\subsection{Defunctionalization}
%specify higher-order definitional interpreters using first-order means (cite)
%\subsection{Refunctionalization}
%\section{Detour: Shiver's \textit{k}-CFA?}

@@

\textbf{Abstracting Abstract Machines.}

\textbf{Pushdown Control-flow Analysis.}
In recent years, there are significant efforts \cite{vardoulakis2010cfa2, earl2012introspective, 
gilray2016pushdown, johnson2015abstracting} to achieve precise call/return 
match based on small-step abstracting abstract machines.

CFA2 is the first solution that solves the return-flows problem\cite{vardoulakis2010cfa2}.
But CFA2 has several drawbacks: it works on only continuation-passing style programs, 
and not supports polyvariant analysis, in addition with an exponential time
complexity.

TODO: summarize PDCFA, AAC

\citeauthor{gilray2016pushdown} proposed a polyvariant continuation-addresses allocator
for small-step AAM to achieve pushdown analysis \cite{gilray2016pushdown}.
The method is both simple to implement and computational inexpensive.
Based on the AAM we presented in Section \ref{aam}, the only change in the code is
not only keeping track of the entry-point expression of callee, but also holding the target environment
when allocating a continuation addresses.
The following code shows the implementation:

@@

The target environment is used for successor states and extended from closure's
environment (namely @val new_env = env_c + (v -> baddr)@).
All the other part of the code would not be changed.
Notably, this change only causes constant-factor increase in time complexity of 
the analysis if the store is widened.

\textbf{Abstracting Definitional Interpreters.}
Reynold's seminal paper \emph{Definitional interpreters for higher-order
  programming languages} \cite{reynolds1972definitional} shows that in
definitional interpreters the defined language can inherit properties from the defining language.
With this insight, \citeauthor{darais2017abstracting} construct abstracting 
definitional interpreters that automatically inherits the pushdown control-flow 
property from its defining language
because the defined language just uses the stack the meta-language \cite{darais2017abstracting}.
\citeauthor{darais2017abstracting}'s abstracting definitional interpreters work
on direct-style programs, and is written in monadic style. One of the advantages
of monadic abstract interpreters is modular and composable, therefore deploying
different sensitivities or features is just applying a different monad.
Prior to that, \citeauthor{Sergey:2013:MAI:2491956.2491979} show a monadic
abstract interpreter for small-step semantics.

This paper is inspired a lot from \citeauthor{darais2017abstracting}'s work.
Instead of just showing the final form of abstract definitional interpreters,
We reveal that refunctionalization plays an important role for inheriting
stack from the meta-language.
The target of the series of transformations in this paper is abstract definitional
interpreters, and we additionaly demonstrate abstract interpreters written in both
monadic- and direct-style.
The difference is ours work for A-Normal Form $\lambda$-calculus instead of plain
$\lambda$-calculus, our work also can be easily adopted to plain
$\lambda$-calculus.
